Response to reviewers

### 2024-11-28 Manuscript/ecography_reponse_letter.tex
updated responses to reviewer 1

\textcolor{blue}{Line 3-6: Please rephrase to simplify and avoid brackets.}
%  We changed the sentence to:
%  \textit{Indeed, functional diversity metrics allow researchers to summarise complex occupancy patterns in space and/or time across communities and/or populations in response to various stressors
% In other words, investigating what, how and why something is changing in an ecosystem by looking at changes of patterns under a certain process through a specific mechanism.}
%  Lines @@@

\textcolor{blue}{Line 10-11: Please rephrase to avoid brackets.}
%  We changed the sentence to:
%  \textit{We studied the ability of different functional diversity metrics to recover patterns and signals from different processes linked to common assembly mechanisms in community ecology; such as environmental filtering, competitive exclusion, equalising fitness, and facilitation.}
%  Lines @@@

\textcolor{blue}{Line 14-15: If space requirements permit, please include more detailed information about your results.
The current description is too vague and does not effectively convey your main findings.}
%  We've expanded a little on the results
% One point of keeping the vagueness in the abstract though was to not give the impression that we provide any prescriptive results but rather a broad understanding that the choice of metric is important.
%  \textit{We show that different metrics perform differently when trying to capture signal from different approximation of common mechanisms relative to no mechanism at all (null).}
%  \textit{For example, the competition was harder to disentangle from the null mechanisms compared to facilitation in our simulations.}

@@

\textcolor{blue}{Line 25-28. It appears to me that you are using the definition of a functional trait from Violle et al.(2007) in Oikos. Please include a citation for this reference if it is the case.}
%  We've added the reference as suggested.

\textcolor{blue}{Line 37. I would avoid using the term ``new trait-focused ecology.'' While it's true that a more ``explicit'' approach to functional ecology has emerged over the past few decades, the foundation of functional analyses is rooted in the field of ecology itself.
For example, as you mentioned, Darwin's finches represent an early functional analysis centered on beak traits.
Although the specific term ``functional traits'' was less commonly used in the past, broader functional analyses were not uncommon, particularly in plant ecology (see, for example, Calow, P 1987 ``Towards a Definition of Functional Ecology.'' Functional Ecology; and Duarte et al., 1995 ``Comparative Functional Plant Ecology: Rationale and Potentials.'' Trends in Ecology \& Evolution).}
%  This is a really good point
% We've removed the mention of ``new''.

\textcolor{blue}{Line 73-74. Please clarify how this is related to the previous question ``what'' and ``how''.}
%  We've expanded the sentence to:
%  \textit{Mechanism description is then about linking the pattern and the process to some biological properties (sometimes referred to as ``rules'\,') and is at the core of answering the biological question at hand.}
%  Lines @@@

\textcolor{blue}{Line 93-96. Please explain why the chosen metric is not appropriate for the proposed example.}
%  We've expanded the sentence to:
%  \textit{In such a scenario, measuring the occupied area (i.e.~\(km^{2}\)) of each population on each island (a pattern - ``what?'\,') will not be the most appropriate way to understand the potential competition between these populations because it might not give information on whether the two populations actually interact in some way - a very likely condition for competition to take place.}
%  Lines @@@

\textcolor{blue}{Line 100. I think it would be beneficial to more clearly address your research question here. Is your goal simply to describe the patterns of functional diversity across different processes, or are you aiming to test how well different metrics capture the effects of various mechanisms? If it's the latter, how did you test this, and what are your expectations in relation to the null model?}
%  We have now clarified what our goal and expectations are in the last paragraph of the introduction:
%  \textit{Our goal is to assess the relative performance of different metrics to capture the patterns of different mechanisms by comparing the score of various metrics under specifically approximated mechanisms and the absence of any specific mechanism (null mechanism).}
%  \textit{We expect that our ability to capture a pattern does not only depends on the choice of metric (what) but also on the process and mechanism at hand (how and why).}
%  Lines @@@

\textcolor{blue}{Line 102. Please refer to Table 1 together with figure 1, to provide more details on the ecological mechanisms.}
%  We now refer to both the table and the figure.

\textcolor{blue}{Line 103-106.Please provide more information on why you chose to use a combination of empirical and simulated data.
From my understanding, your intention was to apply your framework within a more realistic context.
If that is the case, please clarify this point to strengthen your rationale.}
%  We changed the sentence to:
%  \textit{Beyond simulations, we also used an empirical dataset of Hawaiian bird traits and used pre-historical and historical extinction events on Hawaiian islands (process) resulting in a trait space modified by extinctions (the pattern) capturing the effects of species extinction on species trait distributions (the mechanism).}
%  Lines @@@

\textcolor{blue}{Line 106-111.
This content could be more suitable at the start of the discussion section.
Here you could maybe focus on outlining your hypotheses and predictions rather than presenting your conclusions.
By doing so, you establish a clearer context for interpreting your findings and guide the reader through the rationale behind your study.}
%  We have now moved this sentence to the end of the conclusion (Lines @@@) to end the introduction on a clear description of our goals and expectations:

% \textit{Through a simulation exercise, we analysed different patterns (what) across different processes (how) approximating different mechanisms (why) of interaction between organisms: equalisation, filtering, facilitation and competition (Figure \textbackslash ref\{Fig:simulations\} and \textbackslash ref\{Tab:mechanisms\}).}
%  \textit{Our goal is to test the relative performance of different metrics to capture the patterns of different mechanisms by comparing the score of various metrics under specifically approximated mechanisms and the absence of any specific mechanism (null mechanism).}
%  \textit{We expect that our ability to capture a pattern does not only depends on the choice of metric (what) but also on the process and mechanism at hand (how and why).}
%  \textit{To apply our framework to a realistic context, we also used an empirical dataset of Hawaiian bird traits and looked at how anthropogenic pressures have shaped the trait space (mechanism; why) based extinction events on Hawaiian islands before and after the year 1500 (process; how) and how different metrics (pattern; what) can lead to different interpretations of the data.}
%  \textit{We show that, that with a fixed process and mechanisms approximation, the choice of the statistic to describe the pattern (the disparity or functional diversity metrics) has a great impact on the interpretation of the data with the exact same data being sometimes interpreted in opposite ways.}

% Lines @@@

@@

\textcolor{blue}{Line 115. Please provide more details of the elements of your model, as not all readers may be familiar with the Brownian Motion trait model.}
%  We clarified the sentence as follows:
%  \textit{First, we simulated a single random time dependent trait (Brownian Motion) under a model where lineages only speciate (no extinction; i.e.~a pure birth speciation model) until reaching 200 elements in \textbackslash texttt\{R\} \textbackslash citep\{rcore\} using \textbackslash texttt\{treats\} \textbackslash citep\{guillerme2024treats\}.}
%  Lines @@@

\textcolor{blue}{Line 131. Please explain what the distance D is.}
%  We clarified this point as follows:
%  \textit{Increasing density: by removing the required amount of pairs of points with a variable pairwise distance of at least \(D\) (i.e.~removing the \(n\) pairs of points that are at least \(D\) distance away from each other to remove the desired number of elements).}
%  Lines @@@

>>>>

\textcolor{blue}{Line 3-6: Please rephrase to simplify and avoid brackets.}

We changed the sentence to:
\textit{Indeed, functional diversity metrics allow researchers to summarise complex occupancy patterns in space and/or time across communities and/or populations in response to various stressors
In other words, investigating what, how and why something is changing in an ecosystem by looking at changes of patterns under a certain process through a specific mechanism.}
Lines 3-7.

\textcolor{blue}{Line 10-11: Please rephrase to avoid brackets.}

We changed the sentence to:
\textit{We studied the ability of different functional diversity metrics to recover patterns and signals from different processes linked to common assembly mechanisms in community ecology; such as environmental filtering, competitive exclusion, equalising fitness, and facilitation.}
Lines 9-12.

\textcolor{blue}{Line 14-15: If space requirements permit, please include more detailed information about your results.
The current description is too vague and does not effectively convey your main findings.}

We've expanded a little on the results.
One point of keeping the vagueness in the abstract though was to not give the impression that we provide any prescriptive results but rather a broad understanding that the choice of metric is important.

\textit{We show that different metrics perform differently when trying to capture signal from different approximation of common mechanisms relative to no mechanism at all (null).
For example, the competition was harder to disentangle from the null mechanisms compared to facilitation in our simulations.}
Lines 15-18.

@@

\textcolor{blue}{Line 25-28. It appears to me that you are using the definition of a functional trait from Violle et al.(2007) in Oikos. Please include a citation for this reference if it is the case.}

We've added the reference as suggested.

\textcolor{blue}{Line 37. I would avoid using the term ``new trait-focused ecology.'' While it's true that a more ``explicit'' approach to functional ecology has emerged over the past few decades, the foundation of functional analyses is rooted in the field of ecology itself.
For example, as you mentioned, Darwin's finches represent an early functional analysis centered on beak traits.
Although the specific term ``functional traits'' was less commonly used in the past, broader functional analyses were not uncommon, particularly in plant ecology (see, for example, Calow, P 1987 ``Towards a Definition of Functional Ecology.'' Functional Ecology; and Duarte et al., 1995 ``Comparative Functional Plant Ecology: Rationale and Potentials.'' Trends in Ecology \& Evolution).}

This is a really good point
We've removed the mention of ``new''.

\textcolor{blue}{Line 73-74. Please clarify how this is related to the previous question ``what'' and ``how''.}

We've expanded the sentence to:
\textit{Mechanism description is then about linking the pattern and the process to some biological properties (sometimes referred to as ``rules'') and is at the core of answering the biological question at hand.}
Lines 77-79.

\textcolor{blue}{Line 93-96. Please explain why the chosen metric is not appropriate for the proposed example.}

We've expanded the sentence to:
\textit{In such a scenario, measuring the occupied area (i.e. $km^{2}$) of each population on each island (a pattern - ``what?'') will not be the most appropriate way to understand the potential competition between these populations because it might not give information on whether the two populations actually interact in some way - a very likely condition for competition to take place.}
Lines 99-103.

\textcolor{blue}{Line 100. I think it would be beneficial to more clearly address your research question here. Is your goal simply to describe the patterns of functional diversity across different processes, or are you aiming to test how well different metrics capture the effects of various mechanisms? If it's the latter, how did you test this, and what are your expectations in relation to the null model?}

We have now clarified what our goal and expectations are in the last paragraph of the introduction:
\textit{Our goal is to assess the relative performance of different metrics to capture the patterns of different mechanisms by comparing the score of various metrics under specifically approximated mechanisms and the absence of any specific mechanism (null mechanism).
We expect that our ability to capture a pattern does not only depends on the choice of metric (what) but also on the process and mechanism at hand (how and why).}
Lines 109-114.

\textcolor{blue}{Line 102. Please refer to Table 1 together with figure 1, to provide more details on the ecological mechanisms.}

We now refer to both the table and the figure.

\textcolor{blue}{Line 103-106.Please provide more information on why you chose to use a combination of empirical and simulated data.
From my understanding, your intention was to apply your framework within a more realistic context.
If that is the case, please clarify this point to strengthen your rationale.}

We changed the sentence to:
\textit{To apply our framework to a realistic context, we also used an empirical dataset of Hawaiian bird traits and looked at how anthropogenic pressures have shaped the trait space (mechanism; why) based on extinction events on Hawaiian islands before and after the year 1500 (process; how), and how different metrics (pattern; what) can lead to different interpretations of the data..}
Lines 114-119.

\textcolor{blue}{Line 106-111.
This content could be more suitable at the start of the discussion section.
Here you could maybe focus on outlining your hypotheses and predictions rather than presenting your conclusions.
By doing so, you establish a clearer context for interpreting your findings and guide the reader through the rationale behind your study.}

We have now moved this sentence to the end of the conclusion (Lines 569-572) to end the introduction on a clear description of our goals and expectations:

\textit{Through a simulation exercise, we analysed different patterns (what) across different processes (how) approximating different mechanisms (why) of interaction between organisms: equalisation, filtering, facilitation and competition (Figure 1 and Table 1).
Our goal is to test the relative performance of different metrics to capture the patterns of different mechanisms by comparing the score of various metrics under specifically approximated mechanisms and the absence of any specific mechanism (null mechanism).
We expect that our ability to capture a pattern does not only depends on the choice of metric (what) but also on the process and mechanism at hand (how and why).
To apply our framework to a realistic context, we also used an empirical dataset of Hawaiian bird traits and looked at how anthropogenic pressures have shaped the trait space (mechanism; why) based on extinction events on Hawaiian islands before and after the year 1500 (process; how), and how different metrics (pattern; what) can lead to different interpretations of the data.
We show that, that with a fixed process and mechanisms approximation, the choice of the statistic to describe the pattern (the disparity or functional diversity metrics) has a great impact on the interpretation of the data with the exact same data being sometimes interpreted in opposite ways.}
Lines 107-122.

@@

\textcolor{blue}{Line 115. Please provide more details of the elements of your model, as not all readers may be familiar with the Brownian Motion trait model.}

We clarified the sentence as follows:
\textit{First, we simulated a single random time dependent trait (Brownian Motion) under a model where lineages only speciate (no extinction; i.e. a pure birth speciation model) until reaching 200 observations in \texttt{R} (R Core Team, 2024) using \texttt{treats} (Guillerme, 2024).}
Lines 125-127.

\textcolor{blue}{Line 131. Please explain what the distance D is.}

We clarified this point as follows:
\textit{Increasing density: by removing the required amount of pairs of points with a variable pairwise distance of at least $D$.
$D$ is a distance estimated for each trait space to correspond to the minimal Euclidean distance that encompasses at least $n$ pairs of points corresponding to the intensity of the stressor.}
Lines 164-167.

### 2024-11-28 Manuscript/main_ecography_R1.tex
updated responses to reviewer 1

Functional diversity is increasingly used alongside taxonomic diversity to describe populations and communities in ecology.
Indeed, functional diversity metrics allow researchers to summarise complex occupancy patterns in space and/or time across communities and/or populations in response to some stressors.
In other words, investigating what, how and why something is changing in an ecosystem by looking at changes of patterns under a certain process through a specific mechanism. 
However, as the diversity of functional diversity metrics and methods increases, it is often not directly clear which metric is more readily appropriate for which question.
We studied the ability of different functional diversity metrics to recover patterns and signals from different processes linked to common assembly mechanisms in community ecology; such as environmental filtering, competitive exclusion, equalising fitness, and facilitation.
Using both simulated data and an empirical dataset affected by more complex and nuanced mechanisms, we tested the effectiveness of different space occupancy metrics to recover the simulated or empirical changes.
We show that different metrics perform differently when trying to capture signal from different approximation of common mechanisms relative to no mechanism at all (null).
For example competition was harder to disentangle from the null mechanisms compared to facilitation in our simulations.
This emphasises the importance of not using a one-size-fits-all metric.
Instead, researchers should carefully consider and test whether a particular metric will be effective in capturing a pattern of interest.

@@

Through a simulation exercise, we analysed different patterns (what) across different processes (how) approximating different mechanisms (why) of interaction between organisms: equalisation, filtering, facilitation and competition (Figure \ref{Fig:simulations} and Table \ref{Tab:mechanisms}).
Our goal is to assess the relative performance of different metrics in capturing the patterns of different mechanisms by comparing the score of various metrics under specifically approximated mechanisms and the absence of any specific mechanism (null mechanism).
We expect that our ability to capture a pattern does not only depend on the choice of metric (what) but also on the process and mechanism at hand (how and why).
To apply our framework to a realistic context, we also used an empirical dataset of Hawaiian bird traits and looked at how anthropogenic pressures have shaped the trait space (mechanism; why) based on extinction events on Hawaiian islands before and after the year 1500 (process; how), and how different metrics (pattern; what) can lead to different interpretations of the data.
We show that, with a fixed process and mechanisms approximation, the choice of statistic to describe the pattern (the disparity or functional diversity metrics) has a great impact on the interpretation of the data, with the exact same data being sometimes interpreted in opposite ways.

>>>>

Functional diversity is increasingly used alongside taxonomic diversity to describe populations and communities in ecology.
Indeed, functional diversity metrics allow researchers to summarise complex occupancy patterns in space and/or time across communities and/or populations in response to various stressors.
In other words, investigating what, how and why something is changing in an ecosystem by looking at changes of patterns under a certain process through a specific mechanism.
However, as the diversity of functional diversity metrics and methods increases, it is often not directly clear which metric is more readily appropriate for which question.
We studied the ability of different functional diversity metrics to recover patterns and signals from different processes linked to common assembly mechanisms in community ecology; such as environmental filtering, competitive exclusion, equalising fitness, and facilitation.
Using both simulated data and an empirical dataset affected by more complex and nuanced mechanisms, we tested the effectiveness of different space occupancy metrics to recover the simulated or empirical changes.
We show that different metrics perform differently when trying to capture signal from different approximation of common mechanisms relative to no mechanism at all (null).
For example competition was harder to disentangle from the null mechanisms compared to facilitation in our simulations.
This emphasises the importance of not using a one-size-fits-all metric.
Instead, researchers should carefully consider and test whether a particular metric will be effective in capturing a pattern of interest.

@@

Through a simulation exercise, we analysed different patterns (what) across different processes (how) approximating different mechanisms (why) of interaction between organisms: equalisation, filtering, facilitation and competition (Figure \ref{Fig:simulations} and Table \ref{Tab:mechanisms}).
Our goal is to test the relative performance of different metrics to capture the patterns of different mechanisms by comparing the score of various metrics under specifically approximated mechanisms and the absence of any specific mechanism (null mechanism).
We expect that our ability to capture a pattern does not only depends on the choice of metric (what) but also on the process and mechanism at hand (how and why).
To apply our framework to a realistic context, we also used an empirical dataset of Hawaiian bird traits and looked at how anthropogenic pressures have shaped the trait space (mechanism; why) based on extinction events on Hawaiian islands before and after the year 1500 (process; how), and how different metrics (pattern; what) can lead to different interpretations of the data.
We show that, that with a fixed process and mechanisms approximation, the choice of the statistic to describe the pattern (the disparity or functional diversity metrics) has a great impact on the interpretation of the data with the exact same data being sometimes interpreted in opposite ways.
